IML Term project paper

### 2022-12-11 paper.rmd
Naive Bayes.

### Linear Discriminant Analysis (LDA)

A linear discriminant analysis classifier assumes that all $p$ predictors $X = (X_1, \dots, X_p)$ are drawn from a multivariate Gaussian distribution. This means that each predictor follows a normal distribution $N(\mu_k, \Sigma)$ where $\mu_k$ is the class-specific mean and $\operatorname{Cov}(X) = \Sigma$ the covariance matrix of $X$, and there is some correlation between each pair of predictors. The estimate of $p(X|Y)$ is
\begin{equation*}
p(X|Y) = \frac{1}{(2\pi)^{p/2}|\Sigma|^{1/2}} \exp\left( -\frac{1}{2}(x-\mu_k)^T\Sigma^{-1}(x-\mu_k) \right)
\end{equation*}
meaning that the classifier assigns an observation $X$ to to the class for which
\begin{equation*}
\delta_k(x) = x^T\Sigma^{-1}\mu_k-\frac{1}{2}\mu_k^T\Sigma^{-1}\mu_k + \log \pi_k
\end{equation*}
is maximized. 
The covariance matrix being the same for all classes is a defining feature of linear discriminant analysis. The assumption makes the model linear and reduces the number of parameters to estimate. This reduces the flexibility and hence lowers the variance of the model. Lower variance can mean that the model performs well, but if the assumption is not reflected in the data, bias can be high and overall performance suffers due to the bias-variance trade-off.

@@

A Naive Bayes classifier makes no assumptions on the distribution of the observations, and instead assumes that within the *k*th class, the *p* predictors are independent. In other words, the assumption is that there is no association between the predictors, and therefore no joint distribution to consider.

The assumption of independence between predictors is often not realistic, but the model can perform well nonetheless. This is especially true for small data sets, since estimating a joint distribution requires a large amount of data.

>>>>

### Linear Discriminant Analysis

A linear discriminant analysis (LDA) classifier assumes that all $p$ predictors $X = (X_1, \dots, X_p)$ are drawn from a multivariate Gaussian distribution. This means that each predictor follows a normal distribution $N(\mu_k, \Sigma)$ where $\mu_k$ is the class-specific mean and $\operatorname{Cov}(X) = \Sigma$ the covariance matrix of $X$, and there is some correlation between each pair of predictors. The estimate of $p(X|Y)$ is
\begin{equation*}
p(X|Y) = \frac{1}{(2\pi)^{p/2}|\Sigma|^{1/2}} \exp\left( -\frac{1}{2}(x-\mu_k)^T\Sigma^{-1}(x-\mu_k) \right)
\end{equation*}
meaning that the classifier assigns an observation $X$ to to the class for which
\begin{equation*}
\delta_k(x) = x^T\Sigma^{-1}\mu_k-\frac{1}{2}\mu_k^T\Sigma^{-1}\mu_k + \log \pi_k
\end{equation*}
is maximized. 
The covariance matrix being the same for all classes is a defining feature of linear discriminant analysis. The assumption makes the model linear and reduces the number of parameters to estimate. This reduces the flexibility and hence lowers the variance of the model. Lower variance can mean that the model performs well, but if the assumption is not reflected in the data, bias can be high and overall performance suffers due to the bias-variance trade-off.

@@

A naive Bayes classifier makes no assumptions on the distribution of the observations, and instead assumes that within the *k*th class, the *p* predictors are independent. In other words, the assumption is that there is no association between the predictors, and therefore no joint distribution to consider. With this, the posterior probability can be computed as
\begin{equation*}
p(Y=k|X) = \frac{\pi_k p(X_1|Y=k) \cdots p(X_p|Y=k)}{\sum_i \pi_i p(X_1|Y=i) \cdots p(X_p|Y=i)}.
\end{equation*}
In a naive Bayes model, there are a few possible ways to estimate $p(X|Y)$. For quantitative predictors such as ours, we assume that within each class, the *j*th predictor is drawn from a univariate normal distribution, i.e. $X_j|Y = k \approx N (\mu_{jk}, \sigma^2_jk)$.
The assumption of independence between predictors is often not realistic, but the model can perform well nonetheless. This is especially true for small datasets, since estimating a joint distribution requires a large amount of data.
